Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures

### 2012-02-27 ltx/paper_short_2.tex
first final draft added (paper_short_2.tex). lots of new and adjusted images

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities. The technique uses a Dependent Dirichlet Process Mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

We define the task of automated detection and tracking of arbitrary objects in videos to be three things: extraction, localization, and tracking. Extraction is the task of determining the spatiotemporal regions of a video that constitute objects, localization is the task of finding the positions and/or shapes of distinct objects, and tracking is the task of maintaining the identities of the detected objects over time. An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization. Algorithms able to be applied to a variety of objects and video types are particularly desirable. This paper introduces a new framework involving the use of dependent Dirichlet process mixture models. This framework provides a foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research on methods for general detection and tracking of objects tends to focus on one of either extraction, localization, or tracking. Integrating all three tasks in a system for multiple arbitrary objects and diverse video types is not often a primary focus. Accomplishing the three tasks in a cohesive manner has been pursued more recently and has produced algorithms that can, in a fully automated way, accomplish unsupervised detection, localization, and tracking. An example of general detection and tracking is called ``blob tracking'', which typically refers to methods that perform extraction and localization of objects in each frame independently. Tracking involves finding blobs with similar locations and appearances over time. The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}. These difficulties stem from performing detection via segmentation on individual frames.

This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking. This method provides both a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and a general framework for identifying and isolating objects during tracking. Specifically, this work describes how a dependent Dirichlet process mixture model can accomplish these goals. The DDP is first reviewed in the abstract. This model can be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then present our specific model for data obtained via a basic extraction procedure. We describe inference algorithms for the DDP, and specifically our specific model, that allow object localization and tracking to be carried out. We demonstrate our implementation on multiple synthetic and benchmark datasets. Standard performance metrics for each supports our hypothesis that general extraction combined withe sophisticated inference can perform detection and tracking at a level comperable to state-of-the-art, object specific algorithms.

@@

Extraction methods aim to discern the foreground regions of a video from the background. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing pairs of consecutive images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. Another technique for discerning moving regions in an image involves optical flow \citep{horn_1981, bobick_2001}. In \cite{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are found through analysis of optical flow.

@@

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method attempting to carry out this goal, has been used for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects of videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions that can be gained by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and a graph based method that carries out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach based on a Gaussian mixture model has also been used to cluster data gained from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new tracks.

Nearly all methods capable of high accuracy detection and tracking of objects in videos are those tailored for specific object types. These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of objects in a video. The method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos. The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of extraction data to perform detection and tracking. This method begins by performing a simple data extraction procedure (that is consistent for all videos), which can be carried out broadly but yields noisy data; we model this noisy extracted data with a sophisticated type of Bayesian mixture model. Inference can then be performed on this model to gain detection and tracking results.

This paper describes the data extraction procedures (Section~\ref{sec:dataextraction}), the general form of the model (Section~\ref{sec:modeldefinition}), the specific distributions used in our implementation of the model (Section~\ref{sec:modelspecification}), inference procedures used to get tracking results (Section~\ref{sec:inference}), and experimental results (Section~\ref{sec:experiments}).

% Overview
% --------

% \section{Overview of Method}
% \label{sec:overviewofmethod}

@@

\begin{figure*}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_4.pdf}} 
  %../img/longimg_7_detail.pdf}} 
  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames (a - f). The final image shows the result of frame differencing over a sequence of images (from the PETS2009 dataset), where color denotes frame number.}
  \label{fig:img_and_framediff}
\end{figure*}

@@

A basic set of observations that can be extracted from video are three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel. Extraction procedures, such as those discussed in Section~\ref{sec:introduction}, are used to isolate pixels lying in the regions of a video that represent objects. For implementation in experiments (Section~\ref{sec:experiments}), we desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos. Accordingly, we carry out a method called ``frame differencing'' for all experiments, which locates regions within video frames that contain motion by recording the positions of pixels exhibiting large enough differences in intensity or value between succesive frames. Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods useful for non-stationary video, such as \cite{chien2002efficient, zhang2007moving}). Examples of pixel location data extracted via frame differencing are shown in Fig.~\ref{fig:img_and_framediff}(a-j).

Incorporating color information into our model allows us to infer object color, better maintain the identity of adjacent objects, and track objects through occlusion with greater accuracy. We therefore capture the color of each extracted pixel. Specifically, for each pixel $\bold{x} = ( x_{1}, x_{2}, t )$ recorded during frame differencing, we specify a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contains a set of pixels surrounding $\bold{x}$ in frame $t$. We choose a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel. The color value of each of the selected pixels surrounding $\bold{x}$ is recorded. Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of ``color counts''. This vector represents a discrete distribution over colors in the vicinity of a pixel. We will refer to this extraction technique as ``color counting''.

The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
\begin{equation}
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index. 
Additional features may be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations. We leave these attempts for future studies.

@@

@@

A type of Dependent Dirichlet Process Mixture (DDPM) model known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) is used to provide a model-based, time dependent clustering of the extracted data. The model definition is provided in Section~\ref{sec:gpuddpm}, a secondary form of the model used in one of the two inference algorithms is given in Section~\ref{sec:MCMC}, and background information introducing mixture models and the Dirichlet process are given in Appendix~\ref{sec:modelbackground}.

@@

The extracted data contain a collection of spatiotemporal clusters, each representing an object in the video. The number of clusters present in each time step may not be constant (which is to say, clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps), since objects can enter and exit a scene. To cluster data with these properties, we choose a DDPM known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006. This model may be intuitively viewed as a sequence of DPMs, where the parameters and number of clusters are allowed to vary in a specified way between adjacent time steps. Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this application, corresponds to the number of objects---to be found in an automated manner.

As in typical mixture models, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. Usually, the size of each cluster $m_{k,t}$ is a function of the number of observations that have been assigned to it, but in this model, cluster sizes may decrease over time as observations become ``unassigned''; this allows clusters to decrease in strength and potentially die. We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$) conditioned on the previous cluster size ($m_{k,t-1}$), assignments at time $t$ ($c_{1:N_{t}, t}$), and deletion parameter $\rho$ to be
\begin{equation}
\begin{split}
\label{D_distro}
\text{D} &(m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := \hspace{2mm} m_{k,t-1} \\
&- \text{Binomial}( m_{k,t} | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
\end{split}
\end{equation}
$\forall k \in \{1, \ldots, K_{t} \}$, where $K_{t}$ is the number of clusters at time $t$ and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise. We also define a distribution over the assignment of an observation $i$ at time $t$ to a cluster $k$ ($c_{i,t} = k$) conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter $\alpha$, to be
\begin{equation}
\begin{split}
\label{C_distro}
\text{C} ( c_{i,t} = k | &m_{1:K_{t},t}, \hspace{1mm} \alpha)\\ :=
& \begin{cases}
\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
\end{cases}
\end{split}
\end{equation}
% \begin{equation}
% \label{C}
% \begin{split}
% \text{C} &(\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) := \text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \\
% &  \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
% \end{split}
% \end{equation}
% where
% \begin{equation}
% \begin{split}
% P (c_{i,t}=k  | m_{1:K_{t},t})
% = \begin{cases}
% \frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
% \frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
% \end{cases}
% \end{split}
% \end{equation}
$\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$ and we give a newly created cluster the index $K_{t+1}$. Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,
\begin{align}
\label{gpuddpm_def}
\begin{split}
m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{D}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
\theta_{k, t} | \theta_{k, t-1}   &\sim
\begin{cases}
  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
  \mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
\end{cases} \\
c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
\end{split}
\end{align}
where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}. The graphical model associated with this formulation is shown in Fig.~\ref{fig:gpuddpm_gm_1}.
\begin{figure}[h]
        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture. Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}

>>>>

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities. The technique uses a Dependent Dirichlet Process Mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

We define unsupervised detection and tracking of arbitrary objects in videos to be the task of automatically identifying the distinct objects present in a sequence of images and determining the path each object follows over time. Techniques that accomplish this task are useful in many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization. By studying this task, we hope to help make progress towards general machine vision algorithms that can learn the positions, appearances, and number of objects present in any video scene.

We break-down this task into three parts: data extraction, localization, and tracking. Data extraction is the act of extracting features from regions of video-frames that constitute objects, localization is the act of finding the positions and/or shapes of distinct objects, and tracking is the act of maintaining the identities of the detected objects over time. This paper introduces a new framework for carrying out these three actions based on a type of dependent Dirichlet process mixture model. This framework provides a foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research related to general detection and tracking of objects tends to focus on one of either extraction, localization, or tracking. Integrating all three tasks in a system for multiple arbitrary objects and diverse video types is not often a primary focus. A few attempts at accomplishing the three tasks in a cohesive manner have been studied in recent years \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.
%\willie{An example of general detection and tracking, called blob tracking, typically refers to methods that extract data and localize objects in each frame independently. Tracking involves finding blobs with similar locations and appearances over time. The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}. These difficulties stem from performing detection via segmentation on individual frames.}
This paper furthers this line of work by providing a model that gives rise to a number of algorithms to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult, video quality is low, and extraction is noisy---and maintain the isolation of distinct objects during tracking and through occlusion.

We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}). To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for general object tracking or tailored to a specific object type for a given application. In our implementation, we give the specific distributions for representing objects in a general manner (Section~\ref{sec:modelspecification}), which we apply to data extracted via a basic frame differencing procedure. We describe inference algorithms for our model, and show how output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}). Our implementation is demonstrated on multiple synthetic and benchmark datasets, and standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}). We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects. Our results support our hypothesis that simple data extraction combined with inference on our model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.

@@

Methods that discern between the foreground and background regions of a video allow for data to be extracted from the areas in each frame where objects exist. Frame differencing and background subtraction are two such methods. Both record locations that exhibit motion relative to the background. Often, background subtraction refers to methods that compare an image containing targets with an image of the background only or with some model of the background that is learned as the video progresses \cite{piccardi2004background}, while frame differencing refers to methods that compare pairs of consecutive images in a video \cite{zhang2001segmentation}. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.

@@

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method to carry out this goal, has been used for many years, and has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based method that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach based on a Gaussian mixture model has also been used to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.

Nearly all methods capable of high accuracy detection and tracking of objects in videos are those tailored for specific object types. These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of the given objects in a video. The method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos. The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of easily extracted data to perform detection and tracking. This method begins by performing a simple data extraction procedure, which can be carried out broadly but yields noisy data; our model is then applied to this noisy extracted data. Inference is performed on this model to achieve detection and tracking results.

@@

\begin{figure*}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_723.pdf}} \hspace{0.1mm}
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_724.pdf}} \hspace{0.1mm}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame1.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame2.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.328\textwidth]{../img/pets2009_framediff.png}}\\
  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_7.pdf}} 
  %../img/longimg_7_detail.pdf}} 
  \caption{Three pairs of consecutive frames and the results produced by taking the pixel-wise frame difference between each pair (a - i). The final image shows the results of frame differencing over a sequence of images (from the PETS2009/2010 dataset.}
  \label{fig:img_and_framediff}
\end{figure*}

@@

The basic set of observations extracted from video are the three-dimensional points
\begin{equation}
  \centering
  \bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \{ 1, \ldots, T \}
\end{equation}
where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel and $t$ is the time index. These data are extracted by recording the positions of pixels lying in regions that contain objects. For our implementation in experiments (Section~\ref{sec:experiments}), we desire a data extraction procedure that is as unsophisticated as possible, to verify that we can perform accurate inference on noisy data extracted using a broadly applicable procedure. Consequentially, we carry out simple frame differencing (introduced in Section~\ref{sec:introduction}) for all experiments, to find pixels at locations where there is motion. This data extraction procedure is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require data extraction methods useful for non-stationary video, such as \cite{chien2002efficient, zhang2007moving}). Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).

We can extract additional features from videos and incorporate this new information into our model. For example, we can extract color information, and modify our model to allow for object color to be inferred during tracking. Adding this feature allows our method to better distingush between adjacent objects and more accurately track objects through occlusion. For each extracted pixel, we compute a $V$ dimensional vector that represents the color distribution in the pixel's immediate vicinity (details on this vector and how it is computed are given in Section~\ref{sec:dataextractioninexperiments}). We call the components of this vector the ``color counts'' of the pixel. The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
\begin{equation}
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the time index. 
Other features may also be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.

@@

% Model Definition
% --------------

@@

A type of Dependent Dirichlet Process Mixture (DDPM) model known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) is applied to the extracted data. We define the generalized form of this model in Section~\ref{sec:gpuddpm}, and specify the distributions used in our implementation of this model in Section~\ref{sec:modelspecification}. We also define a secondary form of this model in Section~\ref{sec:MCMC}, which is used in one of the two inference algorithms. Background information providing a brief introduction to mixture models and the Dirichlet process is contained in Appendix~\ref{sec:modelbackground}.

@@

The extracted data comprise a collection of spatiotemporal clusters, each corresponding with an object in the video. Since objects can enter and exit a scene, the number of clusters present in each time step may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps). To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn Dependent Dirichlet Process Mixture, introduced by \cite{caron_2007}. This model may be intuitively viewed as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps. Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this application, corresponds to the number of objects in a video---to be found in an automated manner.

Similar to a typical mixture model, in this model, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it. In contrast, cluster sizes in the GPUDDPM may decrease over time as observations become ``unassigned''; this allows clusters to decrease in strength and potentially die. We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$) to be
\begin{equation}
\begin{split}
\label{D_distro}
\text{D} &(m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := \hspace{2mm} m_{k,t-1} \\
&- \text{Binomial}( m_{k,t} | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
\end{split}
\end{equation}
$\forall k \in \{1, \ldots, K_{t} \}$, where $K_{t}$ is the number of clusters at time $t$ and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.

We also define a distribution over the assignment of observation $i$ at time $t$ ($c_{i,t}$), conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter ($\alpha$), to be
\begin{equation}
\begin{split}
\label{C_distro}
\text{C} ( c_{i,t} = k | &m_{1:K_{t},t}, \hspace{1mm} \alpha)\\ :=
& \begin{cases}
\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
\end{cases}
\end{split}
\end{equation}
$\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$ and we give a newly created cluster the index $K_{t+1}$.

Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively to be
\begin{align}
\label{gpuddpm_def}
\begin{split}
m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{D}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
\theta_{k, t} | \theta_{k, t-1}   & \sim
\begin{cases}
  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \leq K_{t} \\ % \in \{ 1, \ldots, K_{t} \} \\
  \mathbb{G}_{0} \hspace{1mm} \text{if} \hspace{1mm} k = K_{t+1}
\end{cases} \\
c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
\end{split}
\end{align}
$\forall$ times $t \in \{1, \ldots, T\}$ and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$, where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}. The graphical model associated with this formulation is shown in Figure~\ref{fig:gpuddpm_gm_1}.
\begin{figure}[h]
        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture. The observations at time $t$, $\bold{x}_{1:N_{t},t}$, and their associated assignments $c_{1:N_{t},t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (likewise for those at time $t-1$).}
\end{figure}
